Syntactically Informed Natural Language Inference

### 2019-03-21 cse517/report/main.tex
v1 done

\paragraph{Compare.} Compare each $(p_i, \bm{H}_i)$ and $(h_j, \bm{P_}j)$ pairs
with another feed-forward neural network $G$:
\begin{align}
    \bm{v}_{1,i} &:= G([p_i, \bm{H}_i])\quad \forall i \in [1,\ldots, \ell_p]\,, \notag \\
    \bm{v}_{2,j} &:= G([h_j, \bm{P}_j])\quad \forall j \in [1,\ldots, \ell_h]\,.
\end{align}

\paragraph{Aggregate.} Aggregate each set of comparison vectors:
\begin{align}
\bm{v}_{1} = \sum_{i=1}^{\ell_p} \bm{v}_{1,i} \qquad\,, \qquad
\bm{v}_{2} = \sum_{j=1}^{\ell_h}  \bm{v}_{2,j}\,.
\end{align}
and make a prediction with a final feed-forward layer $H$:
\begin{align}
    \hat{\bm{y}} = H([\bm{v}_1, \bm{v}_2]).
\end{align}

@@

The minimal span-based neural constituency parser and the deep biaffine attention
model for dependency parsing both encode the input sequence with an LSTM. Using
this LSTM, we obtain
syntactic features $\bm{s}_i$ for each index $i$ of the input. $\bm{s}_i$ is either
the final hidden state $\bm{r}_i$ of the LSTM at index $i$ or a projected
representation of $\bm{r}_i$.

In the initial 
iteration of the network, syntax is not incorporated until the very last feed-forward 
layer. This means that syntax is not considered when calculating attention between 
the premise and the hypothesis. We speculated that by introducing syntax to the 
network prior to calculating attention we would see better results as the syntax 
would help influence the attention calculations and reveal the most important phrases 
of each input statement.

In the ``V3" model, we perform the same initial step of embedding the premise and 
hypothesis and expanding their dimensions (through a fully-connected layer) prior to 
attention. However, instead of calculating our similarity matrix using solely the 
embedded hypothesis and premise vectors as input, we also provide the encoded syntax 
parses of both inputs. The concatenation of these four vectors is used to derive the 
similarity matrix which is then utilized in the attention computations between the 
premise and hypothesis. \jbcomment{Insert V3.jpg here?}

@@

\jbcomment{If we're only using data from the dependency parser we should probably 
remove the section following this comment}
Supplemental variations were done on top of these restructures of the network. One 
of the variations that we introduced was a retrained constituency parser. By 
retraining the constituency parser on the Penn Treebank dataset \citep{marcus1993building} 
we could change the dimensions of the encoded syntax vectors to better fit with our models. 
Additionally, we experimented with training our models using the dependency parser model 
derived by \citet{dozat2016deep} to see if the different types of encoded parses influenced 
the performance of the network.

>>>>

\paragraph{Compare.} Compare each $(p_i, \bm{H}_i)$ and $(h_j, \bm{P_}j)$ pairs
with another feed-forward neural network $G$:
\begin{align}
    \bm{v}^p_i &:= G([p_i, \bm{H}_i])\quad \forall i \in [1,\ldots, \ell_p]\,, \notag \\
    \bm{v}^h_j &:= G([h_j, \bm{P}_j])\quad \forall j \in [1,\ldots, \ell_h]\,.
\end{align}

\paragraph{Aggregate.} Aggregate each set of comparison vectors:
\begin{align}
\bm{v}^p = \sum_{i=1}^{\ell_p} \bm{v}^p_i \qquad\,, \qquad
\bm{v}^h = \sum_{j=1}^{\ell_h}  \bm{v}^h_j\,.
\end{align}
and make a prediction with a final feed-forward layer $H$:
\begin{align}\label{eq:predict}
    \hat{\bm{y}} = H([\bm{v}^p, \bm{v}^h]).
\end{align}

@@

The minimal span-based neural constituency parser and the deep biaffine attention
model for dependency parsing both encode the input sequence with an LSTM. Using
this LSTM, we obtain
syntactic features $\bm{s}^p = \langle s^p_1, \dots, s^p_{\ell_p} \rangle$ and
$\bm{s}^h = \langle s^h_1, \dots, s^h_{\ell_h} \rangle$ from the premise and hypothesis,
respectively. An $s_i$ is either
the final hidden state $\bm{r}_i$ of the LSTM at index $i$ or a projected
representation of $\bm{r}_i$.

\subsection{Syntail Architectures}

We will now describe the different model architectures we experimented with.
In general, as the version number increases the complexity of the model will increase.

\paragraph{v1: Naive late fusion of syntactic features.}
The simplest way to incorporate $\bm{s}$ is to concatenate its final representation to
$\bm{v}^p$ and $\bm{v}^h$ as shown in figure~\ref{fig:v1}.
The prediction step in equation~\ref{eq:predict} then becomes:
\begin{align}
    \hat{\bm{y}} = H([\bm{v}^p, s^p_{\ell_p}, \bm{v}^h, s^h_{\ell_h}]).
\end{align}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/v1.png}
    \caption{Syntail v1 architecture.}
\label{fig:v1}
\end{figure}

\paragraph{v2: Using syntactic features to attend.}
In the ``V3" model, we perform the same initial step of embedding the premise and 
hypothesis and expanding their dimensions (through a fully-connected layer) prior to 
attention. However, instead of calculating our similarity matrix using solely the 
embedded hypothesis and premise vectors as input, we also provide the encoded syntax 
parses of both inputs. The concatenation of these four vectors is used to derive the 
similarity matrix which is then utilized in the attention computations between the 
premise and hypothesis. \jbcomment{Insert V3.jpg here?}

@@

\jbcomment{If we're only using data from the dependency parser we should probably 
remove the section following this comment}
Supplemental variations were done on top of these restructures of the network. One 
of the variations that we introduced was a retrained constituency parser. By 
retraining the constituency parser on the Penn Treebank dataset \citep{marcus1993building} 
we could change the dimensions of the encoded syntax vectors to better fit with our models. 
Additionally, we experimented with training our models using the dependency parser model 
derived by \citet{Dozat2016-gs} to see if the different types of encoded parses influenced 
the performance of the network.
